\propertitle

### 2019-07-11 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-07-11T15:32:36+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

Now consider the related question: What is the activity $\yA_{t}$ of the
full network at some specific time bin $t$? If we don't know the exact
sequence of activities during the recording then we don't know $\yA_{t}$;
if we know the frequency distribution $\yF$ of the activity levels, then
our probability for $\yA_{t}$ is $\yFF_{\yA_{t}}$, out of combinatorial
reasons. But if we are also uncertain about $\yF$, with
probability~\eqref{eq:density_F}, then our probability for $\yA_{t}$ by
marginalization is
\begin{equation}
  \label{eq:probability_AT}
  \pf(\yA_{t} \| \yD \yI) =
  \int \yFF_{\yA_{t}}\;   \pf(\yF \| \yD \yI)\;\di\yF;
\end{equation}
our probability distribution for $\yA_{t}$ is thus simply the mean
$\int \yF \pf(\yF \| \yD \yI)\,\di\yF$.

@@

The post-data belief density~\eqref{eq:density_proportional} has a
maximum-entropy distribution as a mode approximation when $\ycs$ is small,
$\yL$ is large, and $\ycs L$ is small. Its mode in fact will be the
constrained $\yF$ that maximizes the relative entropy $\ysh(\yF;\yR)$. It
should be noted that a pre-data belief different from the entropic
one~\eqref{eq:entropic_prior} will lead to a different mode approximation.
For example, a Dirichlet density would lead to a maximum-entropy
distribution with \emph{Burg's} entropy \citey{burg1975} instead of
Shannon's \citep{jaynes1986d_r1996}.

@@

\begin{figure}[!p]
\centering
\includegraphics[width=\linewidth]{_scripts/MC_plot1000.pdf}%
\\[3em]%
\includegraphics[width=\linewidth]{_scripts/MC_marginal1000.pdf}%
\caption{***}
\label{fig:MC1000}
\end{figure}
% mctests3.nb
% mcfd_6mom_unif_afssd40st0np_cok_tr_N1000_t50_L50_T0.333333333333333_s60000_a15758.0500981316
\begin{figure}[!p]
\centering
\includegraphics[width=\linewidth]{_scripts/MC_plot5000.pdf}%
\\[3em]%
\includegraphics[width=\linewidth]{_scripts/MC_marginal5000.pdf}%
\caption{***}
\label{fig:MC5000}
\end{figure}
% mctests3.nb
% mcfd_6mom_unif_afssd32s_cok_tr_N5000_t180_L50_T0.333333333333333_s25800_a329212306.071189
The properties of the belief density~\eqref{eq:density_proportional} are
shown in \fig~\ref{fig:MC1000} for $N=1\,000$ and in \fig~\ref{fig:MC5000}
for $N=5\,000$, considering the first six factorial moments as our data. It
turns out that that six moments lead to practically the same result as
specifying the full frequency distribution for the sample; they are therefore
informationally sufficient.

The top plots in the figures show the mean
$\int \yF \pf(\yF \| \yD \yI)\,\di\yF$
(\textcolor{mypurpleblue}{\textbf{------}}), which is also our belief about the
activity at any single time bin, \eqn~\eqref{eq:probability_AT}, is clearly
different from the maximum-entropy mode. The latter is therefore not a good
approximation for the full density~\eqref{eq:density_proportional}, which
is quite broad, as can be seen from the ten samples
(\textcolor{mygreen}{------}) drawn from it. This broadness is not
surprising: we're making guesses about thousands of neurons from the
observation of fewer than 100. Yet, the mean and the maximum-entropy mode
are qualitatively similar. In fact, for $N=1\,000$ the second peak of the
mean is even higher than the corresponding maximum-entropy one. The samples
also show that generally two high-frequency regions of activity are
expected, especially for $N=5\,000$.

The bottom plots in the figures show our belief distributions for the
marginal frequency of each activity level, $\pf(\yFF_{\yA} \| \yD \yI)$,
with the averages (\textcolor{mypurpleblue}{\textbf{------}}), medians
(\textcolor{mygreen}{\textbf{--~--~--}}), and 16\% and 84\% quantiles
(\textcolor{myyellow}{$\bm{\cdot \cdot \cdot \cdot}$} corresponding to one
standard deviation for a normal distribution).

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-07-11T15:55:31+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

Now consider the related question: What is the activity $\yA_{t}$ of the
full network at some specific time bin $t$? If we don't know the exact
sequence of activities during the recording then we don't know $\yA_{t}$;
if we know the frequency distribution $\yF$ of the activity levels, then
our probability for $\yA_{t}$ is $\yFF_{\yA_{t}}$, out of combinatorial
reasons. But if we are also uncertain about $\yF$, with
probability~\eqref{eq:density_F}, then our probability for $\yA_{t}$ by
marginalization is
\begin{equation}
  \label{eq:probability_AT}
  \pf(\yA_{t} \| \yD \yI) =
  \int \yFF_{\yA_{t}}\;   \pf(\yF \| \yD \yI)\;\di\yF;
\end{equation}
our probability distribution for $\yA_{t}$ is thus simply the mean
$\int \yF\, \pf(\yF \| \yD \yI)\,\di\yF$.

@@

The post-data belief density~\eqref{eq:density_proportional} has a
maximum-entropy distribution as a mode approximation when $\ycs$ is small,
$\yL$ is large, and $\ycs L$ is small. Its mode in fact will be the
constrained $\yF$ that maximizes the relative entropy $\ysh(\yF;\yR)$. It
should be noted that a pre-data belief different from the entropic
one~\eqref{eq:entropic_prior} will lead to a different mode approximation.
For example, a Dirichlet density would lead to a maximum-entropy
distribution with \emph{Burg's} entropy \citep{burg1975} instead of
Shannon's \citep{jaynes1986d_r1996}.

@@

\begin{figure}[!p]
\centering
\includegraphics[width=\linewidth]{_scripts/MC_plot1000.pdf}%
\\[3em]%
\includegraphics[width=\linewidth]{_scripts/MC_marginal1000.pdf}%
\caption{***}
\label{fig:MC1000}
\end{figure}
% mctests3.nb
% mcfd_6mom_unif_afssd40st0np_cok_tr_N1000_t50_L50_T0.333333333333333_s60000_a15758.0500981316
\begin{figure}[!p]
\centering
\includegraphics[width=\linewidth]{_scripts/MC_plot5000.pdf}%
\\[3em]%
\includegraphics[width=\linewidth]{_scripts/MC_marginal5000.pdf}%
\caption{***}
\label{fig:MC5000}
\end{figure}
% mctests3.nb
% mcfd_6mom_unif_afssd32s_cok_tr_N5000_t180_L50_T0.333333333333333_s25800_a329212306.071189
The properties of the belief density~\eqref{eq:density_proportional} in the
case of our first data set are shown in \fig~\ref{fig:MC1000} for
$N=1\,000$ and in \fig~\ref{fig:MC5000} for $N=5\,000$. Only the first six
factorial moments were considered as our data, but it turns out that they
are informationally sufficient, leading to practically the same result as
specifying the full frequency distribution for the sample.

The top plots in the figures show the mean
$\int \yF\, \pf(\yF \| \yD \yI)\,\di\yF$
(\textcolor{mypurpleblue}{\textbf{------}}), which is also our belief about the
activity at any single time bin, \eqn~\eqref{eq:probability_AT}, is clearly
different from the maximum-entropy mode. The latter is therefore not a good
approximation for the full density~\eqref{eq:density_proportional}, which
is quite broad, as can be seen from the ten samples
(\textcolor{mygreen}{------}) drawn from it. This broadness is not
surprising: we're making guesses about thousands of neurons from the
observation of fewer than 100. Yet, the mean and the maximum-entropy mode
are qualitatively similar. In fact, for $N=1\,000$ the second peak of the
mean is even higher than the corresponding maximum-entropy one. The samples
also show that generally two high-frequency regions of activity are
expected, especially for $N=5\,000$.

The bottom plots in the figures show our belief distributions for the
marginal frequency of each activity level, $\pf(\yFF_{\yA} \| \yD \yI)$,
with the means (\textcolor{mypurpleblue}{\textbf{------}}), medians
(\textcolor{mygreen}{\textbf{--~--~--}}), and 16\% and 84\% quantiles
(\textcolor{myyellow}{$\bm{\cdot \cdot \cdot \cdot}$} corresponding to one
standard deviation for a normal distribution). The coincidence of means
and medians indicate that these marginals are very close to normal
distributions. This happens thanks to the marginalization, which integrates
out $N-2$ quantities. Note, however, that the full distribution for $\yF$
is far from normal, as indicated by the pronounced differences between mode
and mean.

These results caution us against taking maximum-entropy solutions too
literally. The full probabilistic analysis allows us to see more clearly
the extent of our uncertainty and to make better-informed guesses.
